A Formal Verification Framework for Smart Contracts Based on Symbolic Execution

### 2024-01-17 papers/static-checker/ecoop2024/ecoop2024.tex
michelson description

Listing~\ref {lst:simple-program-add} contains a short example of a Michelson program.
\begin{lstlisting}[float=tp,captionpos=b,caption={A Michelson program},label={lst:simple-program-add}]
parameter int;
storage   int;
code { 
(*               pair int int :: [] *)
   DUP; UNPAIR;
(* int :: int :: pair int int :: [] *)
   COMPARE; GE; 
(*       bool :: pair int int :: [] *)
   IF {UNPAIR; SUB} {PUSH string 'Unexpected Pair'; FAILWITH};
(*                        int :: [] *)
   NIL operation; PAIR
}
\end{lstlisting}
This program first duplicates its input (\lstinline|DUP|), decomposes
the pair $P\times S$ into its components (\lstinline|UNPAIR|),
compares the two value on top of the stack (\lstinline|COMPARE|), and
transforms the result into a boolean (\lstinline|GE|). The
\lstinline|IF| instruction pops the boolean and either executes the
first (for true) or second (for false) list of instructions. By now
only the initial pair is on the stack. In the true branch,
\lstinline|UNPAIR| decomposes, and then \lstinline|SUB| subtract the two 
components. In the false branch, the \lstinline|PUSH| instruction pushes
a string message on the stack. The \lstinline|FAILWITH| instruction
abort the contract execution: tokens are consumed, but the storage and
the balance of the contract remain unchanged.

We emerge from the conditional with just the result of the addition on
the stack. The \lstinline|NIL| instruction pushes an empty list and
\lstinline|PAIR| creates the requires pair of type $O^* \times S$.

>>>>

To invoke a Michelson contract on the Tezos blockchain, the caller has
to provide tokens for gas and tokens to add to the contract's balance.
The gas tokens are used account for the cost of running the contract;
if contract execution runs out of gas, it is aborted before it can
perform any visible change. The use of the balance is up to the
discretion of the contract code.

Listing~\ref {lst:simple-program-add} contains a short example of a Michelson program.
\begin{lstlisting}[float=tp,captionpos=b,caption={A Michelson program},label={lst:simple-program-add}]
parameter int;
storage   int;
code { 
(*               pair int int :: [] *)
   DUP; UNPAIR;
(* int :: int :: pair int int :: [] *)
   COMPARE; GE; 
(*       bool :: pair int int :: [] *)
   IF {UNPAIR; SUB} {PUSH string 'Unexpected Pair'; FAILWITH};
(*                        int :: [] *)
   NIL operation; PAIR
}
\end{lstlisting}
\begin{lstlisting}[float=tp,captionpos=b,caption={A Michelson program
with iteration},label={lst:simple-program-iter}]
parameter list int;
storage   int;
code { 
(*               pair (list int) int :: [] *)
   UNPAIR;
(*                   list int :: int :: [] *)
   ITER {ADD};
(*                               int :: [] *)
   NIL operation; PAIR
}
\end{lstlisting}
The code is interspersed with comments that indicate the types of
values that are currently on the stack.
This program first duplicates its input (\lstinline|DUP|), decomposes
the pair $P\times S$ into its components (\lstinline|UNPAIR|),
compares the two values on top of the stack (\lstinline|COMPARE|), and
transforms the result into a boolean (\lstinline|GE|). The
\lstinline|IF| instruction pops the boolean and either executes the
first (for true) or second (for false) list of instructions. By now
only the initial pair is on the stack. In the true branch,
\lstinline|UNPAIR| decomposes, and then \lstinline|SUB| subtracts the two 
components. In the false branch, the \lstinline|PUSH| instruction puts
a string message on the stack. The \lstinline|FAILWITH| instruction
aborts the contract execution: the tokens provided with the invocation
are consumed, but the storage and the balance of the contract remain
unchanged.

After the conditional the stack contains the result of the
subtraction. The \lstinline|NIL| instruction pushes an empty list of
operations and 
\lstinline|PAIR| creates the required pair of type $O^* \times S$.

Michelson has several instructions that perform
iteration. Listing~\ref{lst:simple-program-iter} contains an example
that uses the \lstinline|ITER| instruction to add a list of numbers
(the parameter) to the storage. Like other looping instructions,
\lstinline|ITER| takes a list of instructions that comprise the loop
body (in this case just a single instruction \lstinline|ADD|) and
expects a list to traverse on top of the stack. Essentially,
\lstinline|ITER| runs the loop body for each element in the list,
while allowing full access to the stack: each iteration consumes
one list element and (potentially) transforms the whole stack
while maintaining its shape/type.
